---  
  title: "Course Schedule and Readings"  
---  
  While the topics covered are broad, each week will feature different examples from genetics, ecology, molecular, and evolutionary biology highlighting uses of each individual set of techniques. Each topic will have accompanying readings highlighting a general introduction to a technique (required), and one or more general references or examples. For access to blocked readings, biol609.   
\
To see the code of a lecture, load the slides, then change the file extension in your browser bar to .Rmd to get the code. Or go to [this link](https://github.com/biol609/biol609.github.io/tree/master/lectures) for all of the code.

```{r dates, echo=FALSE, message=FALSE, warning=FALSE}
library(lubridate)
current_date = parse_date_time("2019-01-28", orders="ymd")
datestring <- paste(month(current_date), day(current_date), year(current_date), sep="/")
```    

### Week 1.
`r datestring`   
__Lecture__: [Why Advanced Data Analysis?](lectures/01_intro.html),  [Re-Introduction to Bayes](lectures/02_Bayesian_Intro.html)   
__Reading:__ McElreath Ch. 1-2  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-bayes-2019  


### Week 2.
```{r next_date, echo=FALSE }
current_date <- current_date+ days(7)
datestring <- paste(month(current_date), day(current_date), year(current_date), sep="/")
```
`r datestring`  
__Lecture:__ [Sampling from your Posterior](./lectures/03_posterior_sampling.html), [Linear Bayesian Models](./lectures/04_linear_models.html),   
__Reading:__ McElreath Chapter 3-4  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-linreg-2019    
__Homework:__ [HW1](homework/hw1_sampling_linear.html)

### Week 3.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Multiple Predictor Variables (including Categoricals)](./lectures/14_multiple_predictors.html), [Bayesian Multimodel Inference](./lectures/15_mmi.html)   
__Reading:__ McElreath Chapter 5-6  
__Additional Reading:__ [Aho et al 2013 Ecology](http://izt.ciens.ucv.ve/ecologia/Archivos/ECO_POB%202014/ECOPO2_2014/Aho%20et%20al%202014.pdf), [Gelman et al. 2013 on WAIC and LOO](http://link.springer.com/article/10.1007/s11222-013-9416-2)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-mmi-2019    

### Week 4.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Interaction Effects](./lectures/16_interactions.html), [Markov Chain Monte-Carlo Approaches](./lectures/17_mcmc.html)   
__Reading:__ McElreath Chapter 7-8  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-mcmc_entropy-2019    

### Week 5.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture__: [MaxEnt and GLMs - a review](lectures/09_generalized_linear_models.html), [Bayesian GLMs](./lectures/10_bayesian_glm.html), [GLM lab](lectures/11_bayesian_glm_lab.html)   
__Reading:__ McElreath Chapter 9-10, [fitting Gamma models multiple ways](http://seananderson.ca/2014/04/08/gamma-glms.html), [Gamma hurdle model](http://seananderson.ca/2014/05/18/gamma-hurdle.html)    
__Additional Readings:__ [O'Hara 2009](http://byrneslab.net/classes/biol609/O'Hara_2009_glmm.pdf) through section on GLMs, [O'Hara and Kotze 2010](http://byrneslab.net/classes/biol609/o'hara_and_kotze_do_not_log_transform.pdf), [Wharton and Hui 2011](http://byrneslab.net/classes/biol609/Wharton_Hui_2011_Ecology.pdf),  [Hartig DHARMa vignette](https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html)  
__Files:__  `usethis::use_course("https://biol609.github.io/lectures/glm_data.zip")`  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-glm-2019    


### Week 6.
```{r next_date, echo=FALSE}
```
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Overdispersed Models](./lectures/12_overdisp.html), [Zero Inflated Models](./lectures/13_zinfl.html)   
__Reading:__ McElreath Chapter 11, [Introduction to glmmTMB](https://www.biorxiv.org/content/10.1101/132753v1.abstract), [ver Hoef and Boveng 2007 on overdispersion](http://byrneslab.net/classes/biol609/Ver_Hoef_and_Boveng_qp_neg binomial.pdf), [Roset et al 2006 on Overdispersion](http://byrneslab.net/classes/biol609/Rose_et_al_2006_zero_inflation.pdf)  
__Additional Reading:__ [Zuur chapter on zero inflation](http://byrneslab.net/classes/biol609/zuur_zero_inflation.pdf), [Getting started with glmmTMB](https://cran.r-project.org/web/packages/glmmTMB/vignettes/glmmTMB.pdf), [Troubleshooting with glmmTMB]   
__Files:__ [Fishing Duration CSV](./lectures/data/20/fishing_duration.csv)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-zig_zag-2019    


### Week 7.


```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Random Effects](./lectures/07_random_effects.html), [Mixed Models](./lectures/08_mixed_effects.html)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-ranef-2019  
__Reading:__ [Gelman and Hill Ch. 12](http://byrneslab.net/classes/biol609/gelman_hill_ch_12.pdf) and/or [Zuur on Random Effects](http://byrneslab.net/classes/biol609/Zuur_ch_5.pdf), [A Practical Guide to Generalized Linear Mixed Models](https://www.cell.com/trends/ecology-evolution/fulltext/S0169-5347(09)00019-6)    
__Writings on visualization:__ <a href="http://anythingbutrbitrary.blogspot.com/2012/06/random-regression-coefficients-using.html">Random regression coefficients using lme4</a>, <a href="http://stackoverflow.com/questions/13847936/in-r-plotting-random-effects-from-lmer-lme4-package-using-qqmath-or-dotplot">Making mixed model plots look fancy</a>  
__A Little More Reading:__ <a href="http://jonlefcheck.net/2013/03/13/r2-for-linear-mixed-effects-models/">R<sup>2</sup> for mixed models (from Jon Lefcheck)</a>, [lme4 converge warnings and solutions](https://rstudio-pubs-static.s3.amazonaws.com/33653_57fc7b8e5d484c909b615d8633c01d51.html)  
  __Your One Stop FAQ:__ [Ben Bolker's Mixed Model's FAQ](https://bbolker.github.io/mixedmodels-misc/)  
__Files:__ [Mussels](./lectures/data/07/mussels.csv), [Uneven  Mussels](./lectures/data/07/mussels_ragged.csv), [RIKZ Beach data](./lectures/data/08/rikz.csv), [growth data](./lectures/data/08/nestedGrowth.csv)    
__R Packages:__ lme4, lmerTest, [merTools](https://github.com/jknowles/merTools), RLRsim, cAIC4  
  
### Week 8.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Varying Intercept Mixed Models](./lectures/22_varying_intercept.html) in a Bayesian Context, [Varying Slope Models](./lectures/23a_varslope.html)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-mixed_bayesian-2019   
__Reading:__ McElreath Chapter 12, 13   
__Additional Reading:__ [Gelman and Hill Ch. 12](http://byrneslab.net/classes/biol609/gelman_hill_ch_12.pdf) (see the bit on one model written 5 ways), [Gelman and Hill Ch. 13](http://byrneslab.net/classes/biol609/gelman_hill_ch_13.pdf), [Gelman on multiple comparisons in mixed models](http://byrneslab.net/classes/biol609/Gelman_multcomp.pdf)  

### Week 9.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ Fully Hierarchical Models, [Prediction with Mixed Models](./lectures/23b_prediction.html), Group-Mean Centering and Other Model Structures  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-mixed_bayesian-2019   
__Reading:__ [McNeish and Kelly 2018](./Readings/McNeish_Kelley_PsychMethods_2019.pdf), [Causal Model Structures and Random Effects](http://rpubs.com/jebyrnes/causal_mods), [Poe Lecture Slides](./Readings/2018-11-02_wim_poe_clustered_slides.pdf), [Bell et al. 2018a](./Readings/Bell2018_Article_UnderstandingAndMisunderstandi.pdf)  
__Additional Reading:__ [Bell et al 2018b](./Bell 2018 Qual Quant.pdf)


### Week 10.
```{r next_date, echo=FALSE}
```
`r datestring`   
  __Lecture:__ [Temporal Autocorrelation](lectures/05_timeseries.html), [Spatial Autocorrelation](lectures/06_spatial.html)\
  __Etherpad:__ https://etherpad.wikimedia.org/p/609-autocorrelation-2019  
  __Reading:__ [Zuur et al. on temporal autocorrelation](http://byrneslab.net/classes/biol609/Zuur_ch_6_temporal.pdf), [Why OLS is an Unbiased Estimator for GLS](https://dynamicecology.wordpress.com/2012/09/24/why-ols-estimator-is-an-unbiased-estimator-for-gls/), [Zuur et al. on spatial autocorrelation](http://byrneslab.net/classes/biol609/Zuur_ch_7_space.pdf), [Hawkings on the 8 1/2 deadly sins of spatial data analysis](http://byrneslab.net/classes/biol609/hawkins_2011_spatial.pdf)\
  __Files:__ [Birds](./lectures/data/05/allbirds.csv), [Plankton](./lectures/data/05/planktonMeans.csv), [Boreal Forests](./lectures/data/06/Boreality.txt), [Irish EPA Data](./lectures/data/06/SDI2003.txt)  
  __R Packages:__ nlme   

  
### Week 11. 
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Gaussian Proces Models](./lectures/24_gaussian_process.html) and [Rethinking Alternatives](./lectures/25_beyond_rethinking.html)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-beyond-2019   
__Files:__ [predicted GP script](./lectures/data/gp/predict_gpl2.R)  
__Reading:__ McElreath Chapter 14, [Roberts et al. 2012](http://byrneslab.net/classes/biol609/Roberts_et_al_2012_gaussian_process.pdf)  
__Further reading:__ [R users will now inevitably become Bayesians](https://thinkinator.com/2016/01/12/r-users-will-now-inevitably-become-bayesians/), [rstanarm intro](https://cran.r-project.org/web/packages/rstanarm/vignettes/rstanarm.html), [brms overview](https://cran.r-project.org/web/packages/brms/vignettes/brms_overview.pdf), [stan](http://mc-stan.org), [Introduction to INLA](http://www.r-inla.org/events/bayes2013anintroductiontoinlawithacomparisontojagshandson), [more R-INLA examples](http://www.r-inla.org/examples/volume-1)  
  
### Week 12.  
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ Measurement Error and Missing Data, [Model II and Quantile Regression](./lectures/03_rma_quantile.html)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-alt_regression-2019  
__Reading:__ [Warton et al. SMATR](http://byrneslab.net/classes/biol609/Warton_et_al_2012_MEE.pdf), [Warton et al. 2006 Model II Review](http://byrneslab.net/classes/biol609/Warton_et_al-2006-Biological_Reviews.pdf), [Cade and Noon 2003 on Quantile Regression](http://byrneslab.net/classes/biol609/Cade_and_Noon_2003_Quantile.pdf), [Brennen et al. 2015 quantile regression example](http://byrneslab.net/classes/biol609/Brennen_et_al_2015_quantile.pdf), [Colonescu Ch. 8 on Heteroskedasticity](https://bookdown.org/ccolonescu/RPoE4/heteroskedasticity.html)\
__Files:__ [Elk data](./lectures/data/03/ekk_from_dryad.csv), [Clam data](./lectures/data/04/Clams.txt)  
__R Packages:__ Quantreg  
  
### Week 13.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ Causal Modeling
  
### Week 14.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Intro to Structural Equation Models](https://1drv.ms/p/s!AvOIsObbweOroNIqHiVxra6SIrp4gw), [Implementing Piecewise SEM](https://1drv.ms/f/s!AvOIsObbweOroNIitg7KpsLhgffJrw)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-sem-2019  
__Reading:__ [Grace et al 2012 Ecosphere](http://byrneslab.net/classes/biol609/Grace_et_al-2012-Ecosphere.pdf)  
__R packages:__ piecewiseSEM  

### Week 15.
```{r next_date, echo=FALSE}
```
`r datestring`   
__**Final Work Party Week!**__  
__Final Presentations:__  Friday May 17th  
__Final Papers Due:__ May 21st  
