---  
  title: "Course Schedule and Readings"  
---  
  While the topics covered are broad, each week will feature different examples from genetics, ecology, molecular, and evolutionary biology highlighting uses of each individual set of techniques. Each topic will have accompanying readings highlighting a general introduction to a technique (required), and one or more general references or examples. For access to blocked readings, biol609.   
\
To see the code of a lecture, load the slides, then change the file extension in your browser bar to .Rmd to get the code. Or go to [this link](https://github.com/biol609/biol609.github.io/tree/master/lectures) for all of the code.

```{r dates, echo=FALSE, message=FALSE, warning=FALSE}
library(lubridate)
current_date = parse_date_time("2022-01-23", orders="ymd")
datestring <- paste(month(current_date), day(current_date), year(current_date), sep="/")
```    

### Week 1.
`r datestring`   
__Lecture__: [Why Advanced Data Analysis?](lectures/01_intro.html),  [Introduction to Bayes](lectures/02_Bayesian_Intro.html)   
__Reading:__ McElreath Ch. 1-2  
__Deeper Reading:__ [Bayes' Rules](https://www.bayesrulesbook.com/chapter-2)   
__Even Deeper Reading:__ [Bayesian Basics](https://m-clark.github.io/bayesian-basics/)  
__Tidying Rethinking:__ [Small Worlds and Large Worlds](https://bookdown.org/content/4857/small-worlds-and-large-worlds.html)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-bayes-2024  


### Week 2.
```{r next_date, echo=FALSE }
current_date <- current_date+ days(7)
datestring <- paste(month(current_date), day(current_date), year(current_date), sep="/")
```
`r datestring`  
__Lecture:__ [Sampling from your Posterior](./lectures/03_posterior_sampling.html)     
__Reading:__ McElreath Chapter 3, Sampling the Imaginary   
__Deeper Reading:__ [Approximating the Posterior](https://www.bayesrulesbook.com/chapter-6), [Posterior Inference and Prediction](https://www.bayesrulesbook.com/chapter-8)    
__Even Deeper Reading:__ [Are confidence intervals better termed ”uncertainty intervals”?
](http://www.stat.columbia.edu/~gelman/research/published/uncertainty_intervals.pdf), [Assessing uncertainty in physcial constants](https://www.cmu.edu/epp/people/faculty/research/Fischoff-Henrion-Assessing%20uncertainty%20in%20physical%20constants.pdf)  
__Tidying Rethinking:__ [Sampling from your Posterior](https://bookdown.org/content/4857/sampling-the-imaginary.html)   
__Thursday Problems:__ [sampling](lab/sampling.html)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-posteriors-2024    


### Week 3.
```{r next_date, echo=FALSE }
```
`r datestring`  
__Lecture:__  [Linear Bayesian Models](./lectures/linear_models.html)   
__Reading:__ McElreath Chapter 4  
__Deeper Reading:__ [Simple Normal Regression](https://www.bayesrulesbook.com/chapter-9), [Posterior Inference and Prediction](https://www.bayesrulesbook.com/chapter-8)   
__Even Deeper Reading:__ [Gavin Simpson on B-Splines](https://fromthebottomoftheheap.net/2020/06/03/extrapolating-with-gams/), [Michael Clark's GAM Introduction](https://m-clark.github.io/generalized-additive-models/), [Resources for Learning about GAMs in R](https://github.com/noamross/gam-resources)     
__Tidying Rethinking:__ [Geocentric Models](https://bookdown.org/content/4857/geocentric-models.html)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-linreg-2024    
__Lab:__ [Linear Bayesian Models](lab/linear.html)  

### Week 4.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Multiple Predictor Variables](./lectures/multiple_predictors.html)   
__Reading:__ McElreath Chapter 5, Skim Ch. 6 (we did this in 607)  
__Deeper Reading:__ [Extending the Normal Regression Model](https://www.bayesrulesbook.com/chapter-11)   
__Tidying Rethinking:__ [The Many Variables and Spurious Waffles](https://bookdown.org/content/4857/the-many-variables-the-spurious-waffles.html)   
__Lab:__ [Multiple Predictors with Rethinking](lab/multiple_predictors.html)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-multiple_predictors-2024    


### Week 5.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Bayesian Multimodel Inference](./lectures/mmi.html)   
__Reading:__ McElreath Chapter 7  
__Deeper Reading:__ [Aho et al 2013 Ecology](http://izt.ciens.ucv.ve/ecologia/Archivos/ECO_POB%202014/ECOPO2_2014/Aho%20et%20al%202014.pdf), [Gelman et al. 2013 on WAIC and LOO](http://link.springer.com/article/10.1007/s11222-013-9416-2)  
__Tidying Rethinking:__ [Ulysses Compass](https://bookdown.org/content/4857/ulysses-compass.html)
__Etherpad:__ https://etherpad.wikimedia.org/p/609-mmi-2024    


### Week 6.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Interaction Effects](./lectures/16_interactions.html), [Markov Chain Monte-Carlo Approaches](./lectures/17_mcmc.html)   
__Reading:__ McElreath Chapter 8-9  
__Deeper Reading:__ [MCMC Under the Hood](https://www.bayesrulesbook.com/chapter-7). 
__Etherpad:__ https://etherpad.wikimedia.org/p/609-interactions-2024    

### Week 7.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture__: [MaxEnt and GLMs - a review](lectures/09_generalized_linear_models.html), [Bayesian GLMs](./lectures/10_bayesian_glm.html), [GLM lab](lectures/11_bayesian_glm_lab.html)   
__Reading:__ McElreath Chapter 10-11. 
__Deeper Reading:__ [Poisson and Negative Binomial Regression](https://www.bayesrulesbook.com/chapter-12), [Logistic Regression](https://www.bayesrulesbook.com/chapter-13)  
__Files:__  `usethis::use_course("https://biol609.github.io/lectures/glm_data.zip")`  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-glm-2024    


### Week 8.
```{r next_date, echo=FALSE}
```
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Overdispersed Models](./lectures/12_overdisp.html), [Zero Inflated Models](./lectures/13_zinfl.html)   
__Reading:__ McElreath Chapter 12     
__Further Reading:__ [ver Hoef and Boveng 2007 on overdispersion](http://byrneslab.net/classes/biol609/Ver_Hoef_and_Boveng_qp_neg binomial.pdf), [Roset et al 2006 on Overdispersion](http://byrneslab.net/classes/biol609/Rose_et_al_2006_zero_inflation.pdf), [Zuur chapter on zero inflation](http://byrneslab.net/classes/biol609/zuur_zero_inflation.pdf)
__Etherpad:__ https://etherpad.wikimedia.org/p/609-zig_zag-2024    


### Week 9.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Random Effects](./lectures/07_random_effects.html), [Varying Intercept Mixed Models](./lectures/22_varying_intercept.html)
__Etherpad:__ https://etherpad.wikimedia.org/p/609-ranef-2024  
__Reading:__ McElreath Ch. 13  
__Further Reading:__ [Gelman and Hill Ch. 12](http://byrneslab.net/classes/biol609/gelman_hill_ch_12.pdf) and/or [Zuur on Random Effects](http://byrneslab.net/classes/biol609/Zuur_ch_5.pdf), [A Practical Guide to Generalized Linear Mixed Models](https://www.cell.com/trends/ecology-evolution/fulltext/S0169-5347(09)00019-6)    
__Writings on visualization:__ <a href="http://anythingbutrbitrary.blogspot.com/2012/06/random-regression-coefficients-using.html">Random regression coefficients using lme4</a>, <a href="http://stackoverflow.com/questions/13847936/in-r-plotting-random-effects-from-lmer-lme4-package-using-qqmath-or-dotplot">Making mixed model plots look fancy</a>  
__A Little More Reading:__ <a href="http://jonlefcheck.net/2013/03/13/r2-for-linear-mixed-effects-models/">R<sup>2</sup> for mixed models (from Jon Lefcheck)</a>, [lme4 converge warnings and solutions](https://rstudio-pubs-static.s3.amazonaws.com/33653_57fc7b8e5d484c909b615d8633c01d51.html)  

  
### Week 10.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__  [Varying Slope Models](./lectures/23a_varslope.html)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-mixed_bayesian-2024   
__Reading:__ McElreath Chapter 13   
__Further Reading:__ [Gelman and Hill Ch. 12](http://byrneslab.net/classes/biol609/gelman_hill_ch_12.pdf) (see the bit on one model written 5 ways), [Gelman and Hill Ch. 13](http://byrneslab.net/classes/biol609/gelman_hill_ch_13.pdf), [Gelman on multiple comparisons in mixed models](http://byrneslab.net/classes/biol609/Gelman_multcomp.pdf)  

### Week 11.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__  Instrumental Variables, Group-Mean Centering, and Other Model Structures    
__Reading:__ McElreath Chapter 14 Part 2   
__Etherpad:__ https://etherpad.wikimedia.org/p/609-mixed_bayesian-2024   
__Reading:__ [McNeish and Kelly 2018](./Readings/McNeish_Kelley_PsychMethods_2024.pdf), [Causal Model Structures and Random Effects](http://rpubs.com/jebyrnes/causal_mods), [Poe Lecture Slides](./Readings/2018-11-02_wim_poe_clustered_slides.pdf), [Bell et al. 2018a](./Readings/Bell2018_Article_UnderstandingAndMisunderstandi.pdf), [Bell et al 2018b](./Bell_2018_Qual_Quant.pdf)  


  
### Week 13. 
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Gaussian Proces Models](./lectures/24_gaussian_process.html), [Generalized Additive Models (GAMs)](./lectures/22_gams.html)   
__Reading:__ McElreath Chapter 14 On GP and Ch4 on BSplines   
__Etherpad:__ https://etherpad.wikimedia.org/p/609-beyond-2024   
__Files:__ [predicted GP script](./lectures/data/gp/predict_gpl2.R)  
__Reading:__ [Spatial autoregressive models for statistical inference from ecological data](./Readings/verHoef_spatial_autocorrelation.pdf), McElreath Chapter 14, [Roberts et al. 2012](http://byrneslab.net/classes/biol609/Roberts_et_al_2012_gaussian_process.pdf), [GAMs: an Introduction](http://environmentalcomputing.net/intro-to-gams/), [Michael Clark's GAM Introduction](https://m-clark.github.io/generalized-additive-models/), [Hierarchical Generalized Additive Models in ecology: an introduction with mgcv](./Readings/Pedersen-2024-Hierarchical Generalized Additive Models in ecology.pdf)  
__Further Readings:__ [Methods to account for spatial autocorrelation in the analysis of species distributional data: a review](./Readings/DormannEcography30_57164.pdf)

  
### Week 14.  
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ Measurement Error, Missing Data, and Meta-Analysis
__Reading:__ McElreath Chapter 15.  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-alt_regression-2024  
__Reading:__ [Warton et al. SMATR](http://byrneslab.net/classes/biol609/Warton_et_al_2012_MEE.pdf), [Warton et al. 2006 Model II Review](http://byrneslab.net/classes/biol609/Warton_et_al-2006-Biological_Reviews.pdf), [Cade and Noon 2003 on Quantile Regression](http://byrneslab.net/classes/biol609/Cade_and_Noon_2003_Quantile.pdf), [Brennen et al. 2015 quantile regression example](http://byrneslab.net/classes/biol609/Brennen_et_al_2015_quantile.pdf), [Colonescu Ch. 8 on Heteroskedasticity](https://bookdown.org/ccolonescu/RPoE4/heteroskedasticity.html)\
__Files:__ [Elk data](./lectures/data/03/ekk_from_dryad.csv), [Clam data](./lectures/data/04/Clams.txt)  
__R Packages:__ Quantreg  

### Week 15.
```{r next_date, echo=FALSE}
```
`r datestring`   
__**Final Work Party Week!**__  
__Final Presentations:__  Friday May 17th  
__Final Papers Due:__ May 21st  
