---  
  title: "Course Schedule and Readings"  
---  
  While the topics covered are broad, each week will feature different examples from genetics, ecology, molecular, and evolutionary biology highlighting uses of each individual set of techniques. Each topic will have accompanying readings highlighting a general introduction to a technique (required), and one or more general references or examples. For access to blocked readings, biol609.   
\
To see the code of a lecture, load the slides, then change the file extension in your browser bar to .Rmd to get the code. Or go to [this link](https://github.com/biol609/biol609.github.io/tree/master/lectures) for all of the code.

```{r dates, echo=FALSE, message=FALSE, warning=FALSE}
library(lubridate)
current_date = parse_date_time("2022-01-23", orders="ymd")
datestring <- paste(month(current_date), day(current_date), year(current_date), sep="/")
```    

### Week 1.
`r datestring`   
__Lecture__: [Why Advanced Data Analysis?](lectures/01_intro.html),  [Re-Introduction to Bayes](lectures/02_Bayesian_Intro.html)   
__Reading:__ McElreath Ch. 1-2  
__Deeper Reading:__ [Bayes' Rules](https://www.bayesrulesbook.com/chapter-2)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-bayes-2024  


### Week 2.
```{r next_date, echo=FALSE }
current_date <- current_date+ days(7)
datestring <- paste(month(current_date), day(current_date), year(current_date), sep="/")
```
`r datestring`  
__Lecture:__ [Sampling from your Posterior](./lectures/03_posterior_sampling.html)     
__Reading:__ McElreath Chapter 3.  
__Deeper Reading:__ [Approximating the Posterior](https://www.bayesrulesbook.com/chapter-6), [Bayesian Basics](https://m-clark.github.io/bayesian-basics/), [Posterior Inference and Prediction](https://www.bayesrulesbook.com/chapter-8)    
__Etherpad:__ https://etherpad.wikimedia.org/p/609-posteriors-2024    


### Week 4.
```{r next_date, echo=FALSE }
```
`r datestring`  
__Lecture:__  [Linear Bayesian Models](./lectures/04_linear_models.html),   
__Reading:__ McElreath Chapter 4  
__Deeper Reading:__ [Simple Normal Regression](https://www.bayesrulesbook.com/chapter-9), [Posterior Inference and Prediction](https://www.bayesrulesbook.com/chapter-8)   
__Etherpad:__ https://etherpad.wikimedia.org/p/609-linreg-2024    

### Week 5.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Multiple Predictor Variables](./lectures/14_multiple_predictors.html),   
__Reading:__ McElreath Chapter 5 - 6  
__Deeper Reading:__ [Extending the Normal Regression Model](https://www.bayesrulesbook.com/chapter-11).   
__Etherpad:__ https://etherpad.wikimedia.org/p/609-multiple_predictors-2024    


### Week 6.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Bayesian Multimodel Inference](./lectures/15_mmi.html)   
__Reading:__ McElreath Chapter 7  
__Deeper Reading:__ [Aho et al 2013 Ecology](http://izt.ciens.ucv.ve/ecologia/Archivos/ECO_POB%202014/ECOPO2_2014/Aho%20et%20al%202014.pdf), [Gelman et al. 2013 on WAIC and LOO](http://link.springer.com/article/10.1007/s11222-013-9416-2)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-mmi-2024    


### Week 7.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Interaction Effects](./lectures/16_interactions.html), [Markov Chain Monte-Carlo Approaches](./lectures/17_mcmc.html)   
__Reading:__ McElreath Chapter 8-9  
__Deeper Reading:__ [MCMC Under the Hood](https://www.bayesrulesbook.com/chapter-7). 
__Etherpad:__ https://etherpad.wikimedia.org/p/609-interactions-2024    

### Week 8.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture__: [MaxEnt and GLMs - a review](lectures/09_generalized_linear_models.html), [Bayesian GLMs](./lectures/10_bayesian_glm.html), [GLM lab](lectures/11_bayesian_glm_lab.html)   
__Reading:__ McElreath Chapter 10-11. 
__Deeper Reading:__ [Poisson and Negative Binomial Regression](https://www.bayesrulesbook.com/chapter-12), [Logistic Regression](https://www.bayesrulesbook.com/chapter-13)  
__Files:__  `usethis::use_course("https://biol609.github.io/lectures/glm_data.zip")`  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-glm-2024    


### Week 9.
```{r next_date, echo=FALSE}
```
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Overdispersed Models](./lectures/12_overdisp.html), [Zero Inflated Models](./lectures/13_zinfl.html)   
__Reading:__ McElreath Chapter 11, [Introduction to glmmTMB](https://www.biorxiv.org/content/10.1101/132753v1.abstract), [ver Hoef and Boveng 2007 on overdispersion](http://byrneslab.net/classes/biol609/Ver_Hoef_and_Boveng_qp_neg binomial.pdf), [Roset et al 2006 on Overdispersion](http://byrneslab.net/classes/biol609/Rose_et_al_2006_zero_inflation.pdf)  
__Further Reading:__ [Zuur chapter on zero inflation](http://byrneslab.net/classes/biol609/zuur_zero_inflation.pdf), [Getting started with glmmTMB](https://cran.r-project.org/web/packages/glmmTMB/vignettes/glmmTMB.pdf), [Troubleshooting with glmmTMB]   
__Files:__ [Fishing Duration CSV](./lectures/data/20/fishing_duration.csv)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-zig_zag-2024    


### Week 10.


```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Random Effects](./lectures/07_random_effects.html), [Mixed Models](./lectures/08_mixed_effects.html)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-ranef-2024  
__Reading:__ [Gelman and Hill Ch. 12](http://byrneslab.net/classes/biol609/gelman_hill_ch_12.pdf) and/or [Zuur on Random Effects](http://byrneslab.net/classes/biol609/Zuur_ch_5.pdf), [A Practical Guide to Generalized Linear Mixed Models](https://www.cell.com/trends/ecology-evolution/fulltext/S0169-5347(09)00019-6)    
__Writings on visualization:__ <a href="http://anythingbutrbitrary.blogspot.com/2012/06/random-regression-coefficients-using.html">Random regression coefficients using lme4</a>, <a href="http://stackoverflow.com/questions/13847936/in-r-plotting-random-effects-from-lmer-lme4-package-using-qqmath-or-dotplot">Making mixed model plots look fancy</a>  
__A Little More Reading:__ <a href="http://jonlefcheck.net/2013/03/13/r2-for-linear-mixed-effects-models/">R<sup>2</sup> for mixed models (from Jon Lefcheck)</a>, [lme4 converge warnings and solutions](https://rstudio-pubs-static.s3.amazonaws.com/33653_57fc7b8e5d484c909b615d8633c01d51.html)  
  __Your One Stop FAQ:__ [Ben Bolker's Mixed Model's FAQ](https://bbolker.github.io/mixedmodels-misc/)  
__Files:__ [Mussels](./lectures/data/07/mussels.csv), [Uneven  Mussels](./lectures/data/07/mussels_ragged.csv), [RIKZ Beach data](./lectures/data/08/rikz.csv), [growth data](./lectures/data/08/nestedGrowth.csv)    
__R Packages:__ lme4, lmerTest, [merTools](https://github.com/jknowles/merTools), RLRsim, cAIC4  
  
### Week 8.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Varying Intercept Mixed Models](./lectures/22_varying_intercept.html) in a Bayesian Context, [Varying Slope Models](./lectures/23a_varslope.html)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-mixed_bayesian-2024   
__Reading:__ McElreath Chapter 12, 13   
__Further Reading:__ [Gelman and Hill Ch. 12](http://byrneslab.net/classes/biol609/gelman_hill_ch_12.pdf) (see the bit on one model written 5 ways), [Gelman and Hill Ch. 13](http://byrneslab.net/classes/biol609/gelman_hill_ch_13.pdf), [Gelman on multiple comparisons in mixed models](http://byrneslab.net/classes/biol609/Gelman_multcomp.pdf)  

### Week 9.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__  [brms and Prediction with Mixed Models](./lectures/23c_brms_prediction.html), Group-Mean Centering and Other Model Structures    
__Etherpad:__ https://etherpad.wikimedia.org/p/609-mixed_bayesian-2024   
__Reading:__ [McNeish and Kelly 2018](./Readings/McNeish_Kelley_PsychMethods_2024.pdf), [Causal Model Structures and Random Effects](http://rpubs.com/jebyrnes/causal_mods), [Poe Lecture Slides](./Readings/2018-11-02_wim_poe_clustered_slides.pdf), [Bell et al. 2018a](./Readings/Bell2018_Article_UnderstandingAndMisunderstandi.pdf), [Bell et al 2018b](./Bell_2018_Qual_Quant.pdf)  


### Week 10.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Temporal Autocorrelation](lectures/05_timeseries.html), [Spatial Autocorrelation](lectures/20_spatial.html)\
__Etherpad:__ https://etherpad.wikimedia.org/p/609-autocorrelation-2024  
__Reading:__ [Zuur et al. on temporal autocorrelation](http://byrneslab.net/classes/biol609/Zuur_ch_6_temporal.pdf), [Why OLS is an Unbiased Estimator for GLS](https://dynamicecology.wordpress.com/2012/09/24/why-ols-estimator-is-an-unbiased-estimator-for-gls/), [Zuur et al. on spatial autocorrelation](http://byrneslab.net/classes/biol609/Zuur_ch_7_space.pdf), [Hawkings on the 8 1/2 deadly sins of spatial data analysis](http://byrneslab.net/classes/biol609/hawkins_2011_spatial.pdf)\
__Further Reading:__ [Intro to INLA for spatial and spatiotemporal modeling](https://ourcodingclub.github.io/2018/12/04/inla.html), [Spatial Data Analysis with R-INLA](https://www.jstatsoft.org/article/view/v063i20/v63i20.pdf), [inlabru: a package for Baysianspatial modelling from ecological survey data](https://besjournals.onlinelibrary.wiley.com/doi/abs/10.1111/2041-210X.13168), [Introduction to INLA](http://www.r-inla.org/events/bayes2013anintroductiontoinlawithacomparisontojagshandson), [more R-INLA examples](http://www.r-inla.org/examples/volume-1), [Hierarchical analysis of spatially autocorrelated ecological data using integrated nested Laplace approximation](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/j.2041-210X.2012.00211.x) with [code appendix](https://besjournals.onlinelibrary.wiley.com/action/downloadSupplement?doi=10.1111%2Fj.2041-210X.2012.00211.x&file=MEE3_211_sm_Appendices-data-S1-S2.zip) \
__Files:__ [Birds](./lectures/data/05/allbirds.csv), [Plankton](./lectures/data/05/planktonMeans.csv), [Boreal Forests](./lectures/data/06/Boreality.txt), [Irish EPA Data](./lectures/data/06/SDI2003.txt)\ 
__R Packages:__ nlme   

  
### Week 11. 
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Gaussian Proces Models](./lectures/24_gaussian_process.html), [Generalized Additive Models (GAMs)](./lectures/22_gams.html)   
__Etherpad:__ https://etherpad.wikimedia.org/p/609-beyond-2024   
__Files:__ [predicted GP script](./lectures/data/gp/predict_gpl2.R)  
__Reading:__ [Spatial autoregressive models for statistical inference from ecological data](./Readings/verHoef_spatial_autocorrelation.pdf), McElreath Chapter 14, [Roberts et al. 2012](http://byrneslab.net/classes/biol609/Roberts_et_al_2012_gaussian_process.pdf), [GAMs: an Introduction](http://environmentalcomputing.net/intro-to-gams/), [Michael Clark's GAM Introduction](https://m-clark.github.io/generalized-additive-models/), [Hierarchical Generalized Additive Models in ecology: an introduction with mgcv](./Readings/Pedersen-2024-Hierarchical Generalized Additive Models in ecology.pdf)  
__Further Readings:__ [Methods to account for spatial autocorrelation in the analysis of species distributional data: a review](./Readings/DormannEcography30_57164.pdf)

  
### Week 12.  
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ Measurement Error and Missing Data, [Model II and Quantile Regression](./lectures/03_rma_quantile.html)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-alt_regression-2024  
__Reading:__ [Warton et al. SMATR](http://byrneslab.net/classes/biol609/Warton_et_al_2012_MEE.pdf), [Warton et al. 2006 Model II Review](http://byrneslab.net/classes/biol609/Warton_et_al-2006-Biological_Reviews.pdf), [Cade and Noon 2003 on Quantile Regression](http://byrneslab.net/classes/biol609/Cade_and_Noon_2003_Quantile.pdf), [Brennen et al. 2015 quantile regression example](http://byrneslab.net/classes/biol609/Brennen_et_al_2015_quantile.pdf), [Colonescu Ch. 8 on Heteroskedasticity](https://bookdown.org/ccolonescu/RPoE4/heteroskedasticity.html)\
__Files:__ [Elk data](./lectures/data/03/ekk_from_dryad.csv), [Clam data](./lectures/data/04/Clams.txt)  
__R Packages:__ Quantreg  
  
### Week 13.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ Causal Modeling  
__Readings:__ [A Second Change to Get Causal Inference Right](./Readings/Hernán_and_Healey_2024.pdf), [Causal analysis in control impact studies](./Readings/Larsen_et_al_2024.pdf), [Pearl 2010](Pearl_2010_causal_inference.pdf)    
__Further Reading:__ [Causal Inference](https://www.hsph.harvard.edu/miguel-hernan/causal-inference-book/)
  
### Week 14.
```{r next_date, echo=FALSE}
```
`r datestring`   
<!-- __Lecture:__ [Intro to Structural Equation Models](https://1drv.ms/p/s!AvOIsObbweOroNIqHiVxra6SIrp4gw), [Implementing Piecewise SEM](https://1drv.ms/f/s!AvOIsObbweOroNIitg7KpsLhgffJrw)  -->
__Lecture:__ [Intro to Structural Equation Models](https://www.dropbox.com/s/l7z6sbs16jk2y0s/23_SEM_Intro.pptx?dl=0), [SEM Implementation](https://www.dropbox.com/s/j59oodlmcx53tiw/24_Estimation_in_SEM.pptx?dl=0)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-sem-2024  
__Reading:__ [Grace et al 2012 Ecosphere](http://byrneslab.net/classes/biol609/Grace_et_al-2012-Ecosphere.pdf), [Lefcheck 2016](./Readings/Lefcheck_2016_SEM.pdf)   
___Further Reading:__ [Notes from Jon Lefcheck](https://jslefche.github.io/sem_book/)  
__Helpful Websites:__ [piecewiseSEM](http://jslefche.github.io/piecewiseSEM/), [lavaan](http://lavaan.ugent.be/), [Jim Grace's Tutorials](https://www.usgs.gov/centers/wetland-and-aquatic-research-center/science/quantitative-analysis-using-structural-equation)    

### Week 15.
```{r next_date, echo=FALSE}
```
`r datestring`   
__**Final Work Party Week!**__  
__Final Presentations:__  Friday May 17th  
__Final Papers Due:__ May 21st  
