---  
  title: "Course Schedule and Readings"  
---  
  While the topics covered are broad, each week will feature different examples from genetics, ecology, molecular, and evolutionary biology highlighting uses of each individual set of techniques. Each topic will have accompanying readings highlighting a general introduction to a technique (required), and one or more general references or examples. For access to blocked readings, biol609.   
\
To see the code of a lecture, load the slides, then change the file extension in your browser bar to .Rmd to get the code. Or go to [this link](https://github.com/biol609/biol609.github.io/tree/master/lectures) for all of the code.

```{r dates, echo=FALSE, message=FALSE, warning=FALSE}
library(lubridate)
current_date = parse_date_time("2022-01-23", orders="ymd")
datestring <- paste(month(current_date), day(current_date), year(current_date), sep="/")
```    

### Week 1.
`r datestring`   
__Lecture__: [Why Advanced Data Analysis?](lectures/01_intro.html),  [Introduction to Bayes](lectures/02_Bayesian_Intro.html)   
__Reading:__ McElreath Ch. 1-2  
__Deeper Reading:__ [Bayes' Rules](https://www.bayesrulesbook.com/chapter-2)   
__Even Deeper Reading:__ [Bayesian Basics](https://m-clark.github.io/bayesian-basics/)  
__Tidying Rethinking:__ [Small Worlds and Large Worlds](https://bookdown.org/content/4857/small-worlds-and-large-worlds.html)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-bayes-2024  


### Week 2.
```{r next_date, echo=FALSE }
current_date <- current_date+ days(7)
datestring <- paste(month(current_date), day(current_date), year(current_date), sep="/")
```
`r datestring`  
__Lecture:__ [Sampling from your Posterior](./lectures/03_posterior_sampling.html)     
__Reading:__ McElreath Chapter 3, Sampling the Imaginary   
__Deeper Reading:__ [Approximating the Posterior](https://www.bayesrulesbook.com/chapter-6), [Posterior Inference and Prediction](https://www.bayesrulesbook.com/chapter-8)    
__Even Deeper Reading:__ [Are confidence intervals better termed ”uncertainty intervals”?
](http://www.stat.columbia.edu/~gelman/research/published/uncertainty_intervals.pdf), [Assessing uncertainty in physcial constants](https://www.cmu.edu/epp/people/faculty/research/Fischoff-Henrion-Assessing%20uncertainty%20in%20physical%20constants.pdf)  
__Tidying Rethinking:__ [Sampling from your Posterior](https://bookdown.org/content/4857/sampling-the-imaginary.html)   
__Thursday Problems:__ [sampling](lab/sampling.html)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-posteriors-2024    


### Week 3.
```{r next_date, echo=FALSE }
```
`r datestring`  
__Lecture:__  [Linear Bayesian Models](./lectures/linear_models.html)   
__Reading:__ McElreath Chapter 4  
__Deeper Reading:__ [Simple Normal Regression](https://www.bayesrulesbook.com/chapter-9), [Posterior Inference and Prediction](https://www.bayesrulesbook.com/chapter-8)   
__Even Deeper Reading:__ [Gavin Simpson on B-Splines](https://fromthebottomoftheheap.net/2020/06/03/extrapolating-with-gams/), [Michael Clark's GAM Introduction](https://m-clark.github.io/generalized-additive-models/), [Resources for Learning about GAMs in R](https://github.com/noamross/gam-resources)     
__Tidying Rethinking:__ [Geocentric Models](https://bookdown.org/content/4857/geocentric-models.html)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-linreg-2024    
__Lab:__ [Linear Bayesian Models](lab/linear.html)  

### Week 4.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Multiple Predictor Variables](./lectures/multiple_predictors.html)   
__Reading:__ McElreath Chapter 5, Skim Ch. 6 (we did this in 607)  
__Deeper Reading:__ [Extending the Normal Regression Model](https://www.bayesrulesbook.com/chapter-11)   
__Tidying Rethinking:__ [The Many Variables and Spurious Waffles](https://bookdown.org/content/4857/the-many-variables-the-spurious-waffles.html)   
__Lab:__ [Multiple Predictors with Rethinking](lab/multiple_predictors.html)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-multiple_predictors-2024    


### Week 5.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Bayesian Multimodel Inference](./lectures/mmi.html)   
__Reading:__ McElreath Chapter 7  
__Deeper Reading:__ [Aho et al 2013 Ecology](http://izt.ciens.ucv.ve/ecologia/Archivos/ECO_POB%202014/ECOPO2_2014/Aho%20et%20al%202014.pdf), [Gelman et al. 2013 on WAIC and LOO](http://link.springer.com/article/10.1007/s11222-013-9416-2)  
__Tidying Rethinking:__ [Ulysses Compass](https://bookdown.org/content/4857/ulysses-compass.html)   
__Lab:__ [MultiModel Inference with Rethinking](lab/mmi.html)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-mmi-2024    


### Week 6.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Hamiltonian and Markov Chain Monte-Carlo Approaches](./lectures/mcmc_and_hmc.html)   
__Reading:__ McElreath Chapter 8-9  
__Deeper Reading:__ [MCMC Under the Hood](https://www.bayesrulesbook.com/chapter-7)   
__Lab:__ [HMC and Interaction Effects](lab/mcmc_interaction.html)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-hmc-2024    

### Week 7.
Catch Up Labs!
  
### Week 8.
```{r next_date, echo=FALSE}
```
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture__: [MaxEnt and Bayesian GLMs](lectures/maxent_glm.html)  
__Lab:__ [GLM lab](lectures/11_bayesian_glm_lab.html)   
__Reading:__ McElreath Chapter 10-11. 
__Deeper Reading:__ [Poisson and Negative Binomial Regression](https://www.bayesrulesbook.com/chapter-12), [Logistic Regression](https://www.bayesrulesbook.com/chapter-13)  
__Files:__  `usethis::use_course("https://biol609.github.io/lectures/glm_data.zip")`  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-glm-2024    


### Week 9.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Overdispersed Models](./lectures/12_overdisp.html), [Zero Inflated Models](./lectures/13_zinfl.html)   
__Reading:__ McElreath Chapter 12     
__Further Reading:__ [ver Hoef and Boveng 2007 on overdispersion](http://byrneslab.net/classes/biol609/Ver_Hoef_and_Boveng_qp_neg binomial.pdf), [Roset et al 2006 on Overdispersion](http://byrneslab.net/classes/biol609/Rose_et_al_2006_zero_inflation.pdf), [Zuur chapter on zero inflation](http://byrneslab.net/classes/biol609/zuur_zero_inflation.pdf)
__Etherpad:__ https://etherpad.wikimedia.org/p/609-zig_zag-2024    


### Week 10.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Random Effects](./lectures/07_random_effects.html), [Varying Intercept Mixed Models](./lectures/22_varying_intercept.html),  [Varying Slope Models](./lectures/23a_varslope.html)  
__Etherpad:__ https://etherpad.wikimedia.org/p/609-ranef-2024  
__Reading:__ McElreath Ch. 13, 14 to 14.3  
__Further Reading:__ [Gelman and Hill Ch. 12](http://byrneslab.net/classes/biol609/gelman_hill_ch_12.pdf) and/or [Zuur on Random Effects](http://byrneslab.net/classes/biol609/Zuur_ch_5.pdf), [A Practical Guide to Generalized Linear Mixed Models](https://www.cell.com/trends/ecology-evolution/fulltext/S0169-5347(09)00019-6), [Gelman and Hill Ch. 12](http://byrneslab.net/classes/biol609/gelman_hill_ch_12.pdf) (see the bit on one model written 5 ways), [Gelman and Hill Ch. 13](http://byrneslab.net/classes/biol609/gelman_hill_ch_13.pdf)      
__Writings on visualization:__ <a href="http://anythingbutrbitrary.blogspot.com/2012/06/random-regression-coefficients-using.html">Random regression coefficients using lme4</a>, <a href="http://stackoverflow.com/questions/13847936/in-r-plotting-random-effects-from-lmer-lme4-package-using-qqmath-or-dotplot">Making mixed model plots look fancy</a>  
__A Little More Reading:__ <a href="http://jonlefcheck.net/2013/03/13/r2-for-linear-mixed-effects-models/">R<sup>2</sup> for mixed models (from Jon Lefcheck)</a>, [lme4 converge warnings and solutions](https://rstudio-pubs-static.s3.amazonaws.com/33653_57fc7b8e5d484c909b615d8633c01d51.html)  

  
### Week 11.
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__  Visualizing Complex Bayesian Models    
__Reading:__ TBD

  
### Week 12. 
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ [Gaussian Proces Models](./lectures/24_gaussian_process.html), [Generalized Additive Models (GAMs) and Autocorrelation](./lectures/22_gams.html)   
__Reading:__ McElreath Chapter 14.5 On GP and Ch4 on BSplines   
__Etherpad:__ https://etherpad.wikimedia.org/p/609-beyond-2024   
__Files:__ [predicted GP script](./lectures/data/gp/predict_gpl2.R)  
__Reading:__ [Spatial autoregressive models for statistical inference from ecological data](./Readings/verHoef_spatial_autocorrelation.pdf), McElreath Chapter 14, [Roberts et al. 2012](http://byrneslab.net/classes/biol609/Roberts_et_al_2012_gaussian_process.pdf), [GAMs: an Introduction](http://environmentalcomputing.net/intro-to-gams/), [Michael Clark's GAM Introduction](https://m-clark.github.io/generalized-additive-models/), [Hierarchical Generalized Additive Models in ecology: an introduction with mgcv](./Readings/Pedersen-2024-Hierarchical Generalized Additive Models in ecology.pdf)  
__Further Readings:__ [Methods to account for spatial autocorrelation in the analysis of species distributional data: a review](./Readings/DormannEcography30_57164.pdf)

  
### Week 13.  
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ Measurement Error, Missing Data, and Bayesian Meta-Analysis  
__Reading:__ McElreath Chapter 15.  
__Homework:__ Project Proposals Due  

### Week 14.  
```{r next_date, echo=FALSE}
```
`r datestring`   
__Lecture:__ Bayes Outside of Rethinking  
__Reading:__ brms, INLA, stan?

### Week 15.
```{r next_date, echo=FALSE}
```
`r datestring`   
__**Final Work Party Week!**__  
__Final Presentations:__  Friday May 10th  
__Final Papers Due:__ May 21st  
